---
title: 'Lab 10: ANOVA (by hand)'
author: "JT Larkin"
date: "11/17/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r include=FALSE}
require(here)
rope = read.csv(here("data", "rope.csv"))
rope
```


```{r include=FALSE}
rm(list = ls())

rope = read.csv(here("data", "rope.csv"))
rope$rope.type = factor(rope$rope.type)
levels(rope$rope.type)
       


length(rope$rope.type)
length(levels(rope$rope.type))

summary(rope$rope.type)
n_obs = 121
n_groups = 6

ss_tot = sum((rope$p.cut - mean(rope$p.cut))^2)
ss_tot
df_tot = n_obs - 1
df_tot


aggregate(
  x = rope$p.cut,
  by = list(rope$rope.type),
  FUN = function(x) mean(x))

agg_resids = aggregate(
  x = rope$p.cut,
  by = list(rope$rope.type),
  FUN = function(x) x - mean(x))
str(agg_resids)

agg_sq_resids = aggregate(
  x = rope$p.cut,
  by = list(rope$rope.type),
  FUN = function(x) sum((x - mean(x))^2)) 
str(agg_sq_resids)

ss_within = sum(agg_sq_resids$x)
ss_within
df_within = n_obs - n_groups
df_within

ss_among = ss_tot - ss_within
ss_among
df_among = df_tot - df_within
df_among 

ms_within = ss_among / (n_groups - 1)
ms_within
ms_among  = ss_within / (n_obs - n_groups)
ms_among

f_ratio = ms_among/ms_within
f_ratio
f_pval = pf(f_ratio, df_within, df_among)
f_pval
```






```{r include=FALSE}
#ANOVA in R

fit_1 = lm(p.cut ~ rope.type, data=rope)
anova(fit_1)

anova_fit_1 = anova(fit_1)
str(anova_fit_1)


#anova_fit_1$Sum Sq
anova_fit_1$"Sum Sq"
```


################################################################################

ASSIGNMENT


1. 
```{r}
rm(list = ls())

rope = read.csv(here("data", "rope.csv"))
rope$rope.type = factor(rope$rope.type)
levels(rope$rope.type)
       


length(rope$rope.type)
length(levels(rope$rope.type))

summary(rope$rope.type)
n_obs = 121
n_groups = 6

ss_tot = sum((rope$p.cut - mean(rope$p.cut))^2)
ss_tot
df_tot = n_obs - 1
df_tot


aggregate(
  x = rope$p.cut,
  by = list(rope$rope.type),
  FUN = function(x)mean(x))

agg_resids = aggregate(
  x = rope$p.cut,
  by = list(rope$rope.type),
  FUN = function(x) x - mean(x))
str(agg_resids)

agg_sq_resids = aggregate(
  x = rope$p.cut,
  by = list(rope$rope.type),
  FUN = function(x) sum((x - mean(x))^2)) 
str(agg_sq_resids)

ss_within = sum(agg_sq_resids$x)
ss_within
df_within = n_obs - n_groups
df_within

ss_among = ss_tot - ss_within
ss_among
df_among = df_tot - df_within
df_among 

ms_within = ss_within / (df_within)
ms_within
ms_among  = ss_among / (df_among)
ms_among  
  

f_ratio = ms_among/ms_within
f_ratio
f_pval = 1 - pf(f_ratio, df_among, df_within)
f_pval
```


Self Check
```{r}
# number comparison tolerance
digits_check = 5

# Build the reference model using R functions
fit_1 = lm(p.cut ~ rope.type, data=rope)
anova(fit_1)
anova_fit_1 = anova(fit_1)

# Check degrees of freedom
anova_fit_1$Df == c(df_among, df_within)

# Check sums of squares
round(anova_fit_1$`Sum Sq`, digits = digits_check) == round(c(ss_among, ss_within), digits = digits_check)

# Check mean squares
round(anova_fit_1$`Mean Sq`, digits = digits_check) == round(c(ms_among, ms_within), digits = digits_check)

# Check the F-ratio
round(anova_fit_1$`F value`[1], digits = digits_check) == round(f_ratio, digits = digits_check)

# Check the F test statistic p-value
round(anova_fit_1$`Pr(>F)`[1], digits = digits_check) == round(f_pval, digits = digits_check)
```

2. Looking at the figure I would say that there are not equal variances among groups. I came to this conclusion because the widths of the boxes are not uniform.


3. p-value = 0.00143
```{r}
bartlett.test(agg_resids$x)
```

4. No, looking at the conditional boxplots from the we could tell that there was a potential homogeneity assumption violation. The significant p-value from the Bartlett test confirmed that the variances are not homogeneous and the assumption was violated. Thus, we would need to preform some sort of transformation on the data to deal with the violation.




5. "BLAZE" is the base case rope type.


6.  0.36714. There is no calculation for this value. Since it is the base case, the estimate value is the mean percent cut.

7. 0.2655. We calculate this value by adding the base case estimate value and the "rope.typeXTC" estimate value. So, it would be:  0.36714 + -0.10164 = 0.2655.

```{r}
0.36714 + -0.10164
```

