---
title: 'Lab 10: ANOVA'
author: "Ethan Rutledge"
date: "11/28/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r include=FALSE}
require(here)
rope = read.csv(here("data", "rope.csv"))
rope
```


```{r include=FALSE}
rm(list = ls())

rope = read.csv(here("data", "rope.csv"))
rope$rope.type = factor(rope$rope.type)
levels(rope$rope.type)
       


length(rope$rope.type)
length(levels(rope$rope.type))

summary(rope$rope.type)
n_obs = 121
n_groups = 6

ss_tot = sum((rope$p.cut - mean(rope$p.cut))^2)
ss_tot
df_tot = n_obs - 1
df_tot


aggregate(
  x = rope$p.cut,
  by = list(rope$rope.type),
  FUN = function(x) mean(x))

agg_resids = aggregate(
  x = rope$p.cut,
  by = list(rope$rope.type),
  FUN = function(x) x - mean(x))
str(agg_resids)

agg_sq_resids = aggregate(
  x = rope$p.cut,
  by = list(rope$rope.type),
  FUN = function(x) sum((x - mean(x))^2)) 
str(agg_sq_resids)

ss_within = sum(agg_sq_resids$x)
ss_within
df_within = n_obs - n_groups
df_within

ss_among = ss_tot - ss_within
ss_among
df_among = df_tot - df_within
df_among 

ms_within = ss_among / (n_groups - 1)
ms_within
ms_among  = ss_within / (n_obs - n_groups)
ms_among

f_ratio = ms_among/ms_within
f_ratio
f_pval = pf(f_ratio, df_within, df_among)
f_pval
```






```{r include=FALSE}
#ANOVA in R

fit_1 = lm(p.cut ~ rope.type, data=rope)
anova(fit_1)

anova_fit_1 = anova(fit_1)
str(anova_fit_1)


#anova_fit_1$Sum Sq
anova_fit_1$"Sum Sq"
```




Q1 (8 pts.): Submit the code you used to build your ANOVA by hand. Make sure you use the code template so that you use the same variable names as those which I’ll use for the grading.
```{r}
rm(list = ls())

rope = read.csv(here("data", "rope.csv"))
rope$rope.type = factor(rope$rope.type)
levels(rope$rope.type)
       
length(rope$rope.type)
length(levels(rope$rope.type))

summary(rope$rope.type)
n_obs = 121
n_groups = 6

ss_tot = sum((rope$p.cut - mean(rope$p.cut))^2)
ss_tot
df_tot = n_obs - 1
df_tot

aggregate(x = rope$p.cut,by = list(rope$rope.type), FUN = function(x) mean(x))

agg_resids = aggregate(x = rope$p.cut, by = list(rope$rope.type), FUN = function(x)x - mean(x))

str(agg_resids)

agg_sq_resids = aggregate(x = rope$p.cut,by = list(rope$rope.type),FUN = function(x) sum((x - mean(x))^2)) 

str(agg_sq_resids)

ss_within = sum(agg_sq_resids$x)
df_within = n_obs - n_groups

ss_within
df_within

ss_among = ss_tot - ss_within
df_among = df_tot - df_within

ss_among
df_among 

ms_within = ss_within / (df_within)
ms_among  = ss_among / (df_among)

ms_within
ms_among  
  

f_ratio = ms_among/ms_within
f_pval = 1 - pf(f_ratio, df_among, df_within)

f_ratio
f_pval
```


Self Check: I’ll use the following script to test your answers. You can use it as a self-test prior to submitting your answer.
```{r}
# number comparison tolerance
digits_check = 5

# Build the reference model using R functions
fit_1 = lm(p.cut ~ rope.type, data=rope)
anova(fit_1)
anova_fit_1 = anova(fit_1)

# Check degrees of freedom
anova_fit_1$Df == c(df_among, df_within)

# Check sums of squares
round(anova_fit_1$`Sum Sq`, digits = digits_check) == round(c(ss_among, ss_within), digits = digits_check)

# Check mean squares
round(anova_fit_1$`Mean Sq`, digits = digits_check) == round(c(ms_among, ms_within), digits = digits_check)

# Check the F-ratio
round(anova_fit_1$`F value`[1], digits = digits_check) == round(f_ratio, digits = digits_check)

# Check the F test statistic p-value
round(anova_fit_1$`Pr(>F)`[1], digits = digits_check) == round(f_pval, digits = digits_check)
```

Q2 (1 pt.): Examine the conditional boxplot in the Partitioning Variance: Within-Group section of the walkthrough. Based on the figure, do you think there are equal variances among the groups?

There does not appear to be equal variance among the groups because the sizes of the boxes are different.

Q3 (1 pt.): Conduct a Bartlett test to assess the homogeneity of variances of the percent cut among the rope type groups. Report the p-value.
```{r}
bartlett.test(agg_resids$x)
```
p-value = 0.00143

Q4 (2 pts.): Given your graphical assessment (question 2) and the Bartlett test, do you think an ANOVA-type analysis is appropriate on the raw data? Explain why or why not.

No, the data violates the homogeniety assumption according to the box plots and the significant value in the bartlett test.

```{r}
fit_rope_1 = lm(p.cut ~ rope.type, data = rope)
summary(fit_rope_1)

```

Q5 (1 pt.): Which rope type is the base case?

BLAZE

Q6 (1 pt.): What is the mean percent cut of the base case rope? Show your calculation using value(s) from the model coefficient table.

0.36714, it is what it is because it is the base case there was no calculation.

Q7 (1 pt.): What is the mean percent cut rope type XTC? Show your calculation using value(s) from the model coefficient table.


```{r}
0.36714 + -0.10164
```

