---
title: "Lab 02: R fundamentals 2"
author: "Ethan Rutledge"
date: "9/15/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
n = 12345
vec_1 = sample(12, n, replace = TRUE)
head(vec_1)

vec_2 = vec_1==3

vec_1[vec_2]
```


1. Show the R code you used to create vec_2. :::{.warn} Your code should be a complete and self-contained example. I should be able to paste your code into a fresh R session on my computer and re-create your vec_2 ::


2. Give two reasons why determining which elements in vec_1 have value 3 by visual inspection is a bad idea?
  
  It would most likely result in error and would be time consuming.

```{r}
length(vec_1)
sum(vec_1 == 3)

n = 10
vec_1 = sample(12, n, replace = TRUE)
paste0("Sum of elements with value 3: ", sum(vec_1 == 3))
```

3. Why didn’t you always get the same count of 3 entries each time?

  It is picking a random set of 10 variables between 1 and 12 and           giving you how many of those numbers are 3.So each time you run the       code it is randomized.

4. Considering the different vectors generated each time, explain why using a logical test is a safe way to select entries with a value of 3?

  It proivdes the best random sample, given a specific size of data.        Whereas a human may misrepresent the data by their choice.


5. Explain why performing logical subsetting by hand is very very bad practice. You may want consider re-usability of code, working with different sized data sets, and sharing code with collaborators.

  Using logical subsetting rather than by hand helps discourage human       error and is reproducible, thus can be standardized across                collaborators standardized.


```{r}
for (i in 1:10) {print(paste0("This is loop iteration:", i))}
```


6. Provide the code for your modified loop. It must run as a self-contained example on a fresh R session on my computer.


```{r}
n=5
for (i in 1:n) {print(paste0("This is loop iteration:", i))}
```

7. Provide the code for the modified loop that executes n times. It needs to be a self contained example. I should be able to set the value of n and then run your loop on my computer.


```{r}
n=17
vec_1=sample(1:10, n, replace = TRUE)


for (i in 1:n)
{print(paste0("The element of vec_1 at index", i , " is:" , vec_1[i]))}
```

8. Provide the code you used to create the n, vec_1, and the loop. As always, it should run as a stand-alone example in a fresh R session on my computer.
  Worked with Sonja and JT


```{r}
create_and_print_vec = function(n, min = 1, max =10)
  
{vec_1=sample(min:max, n, replace = TRUE)


for (i in 1:n)
{print(paste0("The element of vec_1 at index", i , " is:" , vec_1[i]))}} 
 
create_and_print_vec(10, min = 100, max = 2000)
```

9. Provide the code you used to build your function.
