---
title: "Data exploration and deterministic functions Lab"
author: "Ethan Rutledge"
date: "9/22/2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: inline
---

```{r}
##install.packages("psych")
require(psych)

```

```{r}
##bring in data and create data frame
require(here)
##read.csv(here("data", "hab.sta.csv"))
dat_habitat = data.frame(read.csv(here("data", "hab.sta.csv")))
head(dat_habitat)

##read.csv(here("data", "bird.sta.csv"))
dat_bird = data.frame(read.csv(here("data", "bird.sta.csv")))
head(dat_bird)
```

```{r, echo=TRUE, width=1600, height=1400}
##merge 2 data sets with identical variables and change image output dimensions
dat_all = merge(dat_habitat, dat_bird)
plot(ba.tot ~ elev, data = dat_all)

```

```{r}
##random sample of 100 variables to check out data
sample(dat_all$CEWA, 100)
```

```{r}
##create boolean vector of  all the times cedar wax wing was observed
cewa_vec = (dat_bird$CEWA > 0)
##change to numeric present absent data and plot it
cewa_present_absent = as.numeric(cewa_vec > 0)
plot(x = dat_all$elev, y = cewa_present_absent)
```

```{r}
# Function to calculate the logistic parameter a given the slope and midpoint
get_logistic_param_a = function(slope, midpoint)
{
  b = slope / 4
  return (-midpoint * (slope / 4))
}

# Function to calculate the logistic parameter b given the slope
get_logistic_param_b = function(slope)
{
  return (slope / 4)
}


# Calculate the value of the logistic function at x, given the parameters a and b.
logistic = function(x, a, b)
{
  val = exp(a + b * x)
  return(val / (1 + val))
}

# Calculate the value of the logistic function at x, given a slope and midpoint.
logistic_midpoint_slope = function(x, midpoint, slope)
{
  b = get_logistic_param_b(slope)
  a = get_logistic_param_a(slope, midpoint)
  return(logistic(x, a, b))}
```

```{r}
##plot a logistic curve. It looks like a good midpoint value could be 400. We can try a slope of 0.1 to start with
plot(x = dat_all$elev, y = cewa_present_absent)
curve(logistic_midpoint_slope(x, midpoint = 400, slope = 0.1), add = TRUE)
```
```{r}
##try a negative slope
plot(x = dat_all$elev, y = cewa_present_absent)
curve(logistic_midpoint_slope(x, midpoint = 400, slope = -0.1), add = TRUE)

```
```{r}
##try a shallower negative slope
plot(x = dat_all$elev, y = cewa_present_absent)
curve(logistic_midpoint_slope(x, midpoint = 400, slope = -0.05), add = TRUE)

```
```{r}
## make it a pretty plot
plot(x = dat_all$elev, y = cewa_present_absent, col = "blue", cex = 1, pch = 16)
curve(logistic_midpoint_slope(x, midpoint = 400, slope = -0.05), add = TRUE)

```
Assignment:

1. Use the pair plot function from psych to create a pair plot of the three terrain variables and basal area from the lecture questions.



```{r}
## create a pair plot using the 3 terrain variables and basal area, the [,c] refers to using a subset from the dat_all dataset from columns elev,slope,aspect, as opposed to [c,] which refers to all rows.
pairs.panels(dat_all[,c("elev", "slope", "aspect", "ba.tot")])

```
2. Choose two bird species and create plots of presence/absence (on the y-axis) and basal area (on the x axes).

3. Visually inspect the plots and fit logistic curves using the parameterization functions provided above.

```{r}
##create boolean vector of  all the times common yellow throat and dove was observed
coye_vec = (dat_all$COYE > 0)
##change to numeric present absent data
coye_present_absent = as.numeric(coye_vec > 0)
##plot it
plot(x = dat_all$ba.tot, y = coye_present_absent, ylab = "Common Yellow Throat Presence/Absence", xlab = "Total Basal Area", main = "Presence of Common Yellow Throat by Total Basal Area")
##fit logistic curve
curve(logistic_midpoint_slope(x, midpoint = 100, slope = -0.5), add = TRUE)
##at very low total basal area we find a few birds, but any thicker and theyre out.

```

```{r}
amro_vec = (dat_all$AMRO > 0)
amro_present_absent = as.numeric(amro_vec > 0)
plot(x = dat_all$ba.tot, y = amro_present_absent)

plot(x = dat_all$ba.tot, y = amro_present_absent, ylab = "American Robin Presence/Absence", xlab = "Total Basal Area")
curve(logistic_midpoint_slope(x, midpoint = 100, slope = -0.5), add = TRUE)
##we see the presence of robins matches the absence of them up to 100m per ha, looks like a 50% chance of seeing them up to that point.

```
Questions:

1. What is basal area, and how is it measured?
Basal area is the amount of area covered by tree trunks. It is measured using a prism and dbh. It's units is m^2 per hectare.

2. Include a figure of your terrain/basal area pairplots
```{r}
dat_all = merge(dat_habitat, dat_bird)
pairs.panels(dat_all[,c("elev", "slope", "aspect", "ba.tot")])
```

3. Include a figure of your logistic function plot. Your figure must include the name of the bird species, appropriate title, axes, etc.
```{r}
coye_vec = (dat_all$COYE > 0)
coye_present_absent = as.numeric(coye_vec > 0)
plot(x = dat_all$ba.tot, y = coye_present_absent, ylab = "Common Yellow Throat Presence/Absence", xlab = "Total Basal Area", main = "Presence of Common Yellow Throat by Total Basal Area")
curve(logistic_midpoint_slope(x, midpoint = 25, slope = -0.5), add = TRUE)
```

4. Qualitatively describe the bird’s presence/absence patterns in terms of basal area.
Your answer should make reference to your fitted logistic model.
Does the bird species seem to prefer areas with high or low tree cover?
Does a logistic model seem like a good fit?

At very low total basal area we find a few birds, probably 1-10 m^2 per ha but any thicker forested areas and they're out. The logistic curve drops off quickly after the few detections, however it still captures a lot of the absences, so it may not be a good fit. I may have done it wrong or am interpreting it wrong.There were very little observations of Yellow Throat in this study so its hard to fit a curve either way.

5. Include a figure of your logistic function plot. Your figure must include the name of the bird species, appropriate title, axes, etc.
```{r}
amro_vec = (dat_all$AMRO > 0)
amro_present_absent = as.numeric(amro_vec > 0)
plot(x = dat_all$ba.tot, y = amro_present_absent, ylab = "American Robin Presence/Absence", xlab = "Total Basal Area", main = "American Robin Presence by Total Basal Area")
curve(logistic_midpoint_slope(x, midpoint = 100, slope = -0.5), add = TRUE)

```

6. Qualitatively describe the bird’s presence/absence patterns in terms of basal area.
Your answer should make reference to your fitted logistic model.
Does the bird species seem to prefer areas with high or low tree cover?
Does a logistic model seem like a good fit?

We see the presence of robins matches the absence of them up to 100m per ha, looks like a 50% chance of seeing them up to that point. This makes sense with how common robins are.The logistic curve makes sense, there is a high probability of seeing robins up to 100 m^2 per ha then it drops off rapidly.

7 and 8. How many total number of Gray Jays were observed in all of the sampling sites.
Try using the dat_all data frame to calculate the answer. Include the R code you used to perform the calculation.
```{r}
sum(dat_all$GRJA)
```
9 and 10. Calculate the total number of sampling sites in which Gray Jays were observed.
Hint: What happens when you use the sum() function on a vector of Boolean values?

```{r}
grja_vec = (dat_all$GRJA > 0)
grja_present_absent = as.numeric(grja_vec > 0)
sum(grja_present_absent)
```































