---
title: "Lab_4_Uncertainty_and_Error"
author: "Ethan Rutledge"
date: "9/29/2021"
output: pdf_document
---

1.Show the code you used to create your vectors.

```{r}
norm_17 = rnorm(n = 17, mean = 10.4, sd = 2.4)
norm_30 = rnorm(n = 30, mean = 10.4, sd = 2.4)
norm_300 = rnorm(n = 300, mean = 10.4, sd = 2.4)
norm_3000 = rnorm(n = 3000, mean = 10.4, sd = 2.4)
```

2.Include the R code you used to create your figure. Your answer should include code that builds the figure as well as saves it to a file.

```{r echo=TRUE, height=1600, width=1500, res=180}

require(here)
png(filename = here("Images", "lab_04_hist_01.png"), width = 1500, height = 1600, 
  res = 180)
par(mfrow = c(2, 2))
hist(norm_17, main = " Normal Distribution of Random Numbers \nUsing 17 Data Points", 
     ylab = "Frequency", xlab = "Random 17 points with a mean \nof 10.4 and st. dev of 2.4")
hist(norm_30, main = " Normal Distribution of Random Numbers \nUsing 30 Data Points", 
     ylab = "Frequency", xlab = "Random 30 points with a mean \nof 10.4 and st. dev of 2.4")
hist(norm_300, main = " Normal Distribution of Random Numbers \nUsing 300 Data Points", 
     ylab = "Frequency", xlab = "Random 300 points with a mean \nof 10.4 and st. dev of 2.4")
hist(norm_3000, main = " Normal Distribution of Random Numbers \nUsing 3000 Data Points", 
     ylab = "Frequency", xlab = "Random 3000 points with a mean \nof 10.4 and st. dev of 2.4")

dev.off()

```

3. Upload your lab_04_hist_01.png file to Moodle.

4. Qualitatively describe the differences among the three histograms.

  There are less bars and more sporadic behavior in the histograms with less data points. 
  There is no clear pattern in the first two histograms.

5. Explain why the shapes of the histograms are different.

  As the histograms use more data points the shapes look closer to a normal 
  distribution/bell shape. This is because with more data points you have a better 
  chance of having more numbers that fall closer to the true mean.

6. What are the parameters and their values for the standard Normal distribution?

  They fall within -1 and 1, with a mean of 0 and a standard deviation of 1.
  
7. Include the R code you used to create your figure. Your answer should include 
code that builds the figure as well as saves it to a file.

```{r echo=TRUE, height=1200, width=1500, res=180}
svg(filename = here("Images", "norm_1.svg"))

x = seq(-20, 20, length.out = 1000)
y = dnorm(x, mean = 10.4, sd=2.4)

plot(x, y, main = "Normal Probability Density Function", type = "l", xlim = c(-20, 20),
     ylab = "Fequency of X", 
     xlab = "Random Numbers [-20~20] with mean = 10.4 \nand St.D = 2.4 using 1000 data points" )
abline(h = 0)

dev.off()
```


8. Upload figure to moodle.


9. Show the R code you used to create one of the random datasets in your figure.


```{r}

png(filename = here("Images", "lab_04_plethira.png"))
par(mfrow = c(2, 2))

n_pts = 500
x_min = 1
x_max = 500
x = runif(n = n_pts, min = x_min, max = x_max)

dat = data.frame(x = x, y_observed = rnorm(n_pts))

plot(y_observed ~ x, data = dat, pch = 20, col = "blue", main = "Random 100 Blue Points", 
     ylab = "Normal Distribution of X Using 500 Data Points", 
     xlab = "Random Points Between 1 and 500")


n_pts_1 = 1000
x_min_1 = 1
x_max_1 = 1000
x_1 = runif(n = n_pts_1, min = x_min_1, max = x_max_1)

dat_1 = data.frame(x = x_1, y_observed_1 = rnorm(n_pts_1))

plot(y_observed_1 ~ x_1, data = dat_1, pch = 20, col = "hotpink", 
     main = "Random 1000 Hot Pink Points", 
     ylab = "Normal Distribution of X Using 1000 Data Points", 
     xlab = "Random Points Between 1 and 1000")


norm_new = rnorm(n = 5000, mean = 2500, sd = 250)

hist(norm_new, main = " Histogram of Random Numbers", col = "lawngreen", 
     ylab = "Frequency", 
     xlab = "Random 5000 Points with a mean = 2500 and st.dev = 250")

boxplot(x_1, x, ylab = "Normal Random Number Distribtuion", col = "coral", 
        main = " Boxplots of Two Random \nNumber Distributions")

dev.off()

```


10. Upload an image file of your figure. It may be in png (raster graphics), svg 
(vector graphics), or pdf format.


11. Show the R code you used to create one of the random datasets in your figure.

```{r echo=TRUE, height=1200, width=2000, res=180}

png(filename = here("Images", "lab_04_Curves.png"))

line_point_slope = function(x, x1, y1, slope)
{
  get_y_intercept = 
    function(x1, y1, slope) 
      return(-(x1 * slope) + y1)
  
  linear = 
    function(x, yint, slope) 
      return(yint + x * slope)
  
  return(linear(x, get_y_intercept(x1, y1, slope), slope))
}



n_pts_1 = 1000
x_min_1 = 1
x_max_1 = 500
x_1 = runif(n = n_pts_1, min = x_min_1, max = x_max_1)

dat_1 = data.frame(x = x_1, y_observed_1 = rnorm(n_pts_1))

Ques_11_x = 250
Ques_11_y = 0
Ques_11_slope = 0.001

plot(y_observed_1 ~ x_1, data = dat_1, pch = 20, col = "hotpink", 
     main = "Random 1000 Hot Pink Points", 
     ylab = "Random Distribution Using 1000 Data Points", 
     xlab = "Random Numbers Between 1 and 500")
curve(line_point_slope(x, Ques_11_x, Ques_11_y, Ques_11_slope), add = T)

dev.off()

```


12. Upload the Image to moodle

13/14. Paste the R code you used to create create the columns of predicted values 
and residuals. Include images.

```{r}
n_pts_1 = 1000
x_min_1 = 1
x_max_1 = 500
x_1 = runif(n = n_pts_1, min = x_min_1, max = x_max_1)

dat_1 = data.frame(x = x_1, y_observed_1 = rnorm(n_pts_1))

y_observed_1 = rnorm(n_pts_1)

y_predicted = line_point_slope(dat_1$x, Ques_11_x, Ques_11_y, Ques_11_slope)

resids = y_observed_1 - y_predicted

Quess_13_df = data.frame(y_observed_1, y_predicted, resids)

plot(resids ~ y_predicted, data = dat_1, pch = 20, col = "blue", 
     main = "Residual Scatterplot", 
     ylab = "Model Residuals", 
     xlab = "Predicted Values")

hist(resids, main = "Histogram of Model Residuals", ylab = "Frequency", 
     xlab = "Model Residuals")

```









