---
title: "Lab_5_Uncertainty, Samples, and Populations"
author: "Ethan Rutledge"
date: "10/13/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Q1. Show the R code you used to create exp_fun()

Q2. In your lab report, include a single figure containing four negative exponential curves with these parameter values
```{r}
exp_fun = function(x, a, b){return(a * exp(-b * x))}

curve(
  exp_fun(x, 1.9, 0.9),
  from = 0, to = 10, add = FALSE, 
  main = "Ethan's Exponential Function",
  ylab = "frequency of (x)", xlab = "x", lty = "solid", col = "black")

curve(
  exp_fun(x, 1.9, 0.3),
  from = 0, to = 10, add = TRUE, 
  main = "Ethan's Exponential Function",
  ylab = "f(x)", xlab = "x", lty = "dotted", col = "black")

curve(
  exp_fun(x, 1.2, 0.2),
  from = 0, to = 10, add = TRUE, 
  main = "Ethan's Exponential Function",
  ylab = "f(x)", xlab = "x", lty = "solid", col = "red")

curve(
  exp_fun(x, 1.2, 0.4),
  from = 0, to = 10, add = TRUE, 
  main = "Ethan's Exponential Function",
  ylab = "f(x)", xlab = "x", lty = "dotted", col = "red")
```
Q3. Observe how the curves vary as you change the two parameters’ values. Qualitatively describe what happens to the curve as you vary parameter a.

  The location where the line starts on the Y axis changes depending on the a value.

Q4. Observe how the curves vary as you change the two parameters’ values. Qualitatively describe what happens to the curve as you vary parameter b.

  The curve seems to be more deep and drastic with larger values of b.

Q5. In your lab report, include a single plot containing 6 Ricker curves with these parameter values:
```{r}
ricker_fun = function(x, a, b) 
{
  return(a * x * exp(-b * x))
}

curve(
  ricker_fun(x, 25, 0.1), 
  from = 0, to = 100, add = FALSE, 
  main = "Ethan's Ricker Function",
  ylab = "frequency of (x)", xlab = "x", lty = "solid", col = "black")

curve(
  ricker_fun(x, 20, 0.2), 
  from = 0, to = 100, add = TRUE, 
  main = "Ethan's Ricker Function",
  ylab = "f(x)", xlab = "x", lty = "dotted", col = "black")

curve(
  ricker_fun(x, 10, 0.2), 
  from = 0, to = 100, add = TRUE, 
  main = "Ethan's Ricker Function",
  ylab = "f(x)", xlab = "x", lty = "dotted", col = "black")

curve(
  ricker_fun(x, 75, 0.3), 
  from = 0, to = 100, add = TRUE, 
  main = "Ethan's Ricker Function",
  ylab = "f(x)", xlab = "x", lty = "solid", col = "red")

curve(
  ricker_fun(x, 50, 0.3), 
  from = 0, to = 100, add = TRUE, 
  main = "Ethan's Ricker Function",
  ylab = "f(x)", xlab = "x", lty = "dotted", col = "red")

curve(
  ricker_fun(x, 40, 0.3), 
  from = 0, to = 100, add = TRUE, 
  main = "Ethan's Ricker Function",
  ylab = "frequency of (x)", xlab = "x", lty = "dotted", col = "red")

```
Q6. Observe how the curves vary as you change the two parameters’ values. Qualitatively describe what happens to the curve as you vary parameter a

  The height of the curve goes up or down with different values of a. Higher curve for higher values of a.

Q7. Observe how the curves vary as you change the two parameters’ values. Qualitatively describe what happens to the curve as you vary parameter b

  It's really hard to tell whats going on when you change the b variable. It seems to only have a lot of weight with varying a values.

Q8. Linear Model. Provide the values of the slope, x1, and y1 parameters you chose. Briefly describe how you chose the values.

  slope= -0.0005 x = 800 y = 0.2 I chose these values because they seemed to be where center of the data was. I chose the slope by trying different numbers until it matched the shape of the data.

Q9. In your lab report, include a scatterplot of the salamander data with your fitted linear model.
```{r}
require(here)

line_point_slope = function(x, x1, y1, slope)
{
  get_y_intercept = 
    function(x1, y1, slope) 
      return(-(x1 * slope) + y1)
  
  linear = 
    function(x, yint, slope) 
      return(yint + x * slope)
  
  return(linear(x, get_y_intercept(x1, y1, slope), slope))
}

sal_dispersal = data.frame(read.csv(here("data", "dispersal.csv")))

plot(disp.rate.ftb ~ dist.class, data = sal_dispersal, 
     main = "Linear Function of Salamander Dispersal", 
     ylab = "Dispersal Rate of FTB", xlab = "Distance Class", pch = 10, col = "Black")

curve(line_point_slope(x, 800, 0.2, -0.0005), add = T)

```

Q10 (2 pts): Exponential Model. Provide the values of the a and b. Briefly describe how you chose the values.

  a = 1.5 and b = 0.003 I chose 1.5 for my a value because I wanted my line to cross the y axis around that area. I played around with the b value until it fit my data.

Q11 (2 pts): In your lab report, include a scatterplot of the salamander data with your fitted exponential model.
```{r}
exp_fun = function(x, a, b){return(a * exp(-b * x))}

plot(disp.rate.ftb ~ dist.class, data = sal_dispersal, pch = 20, col = "hotpink", 
     main = "Exponential Function of Salamander Dispersal", 
     ylab = "Dispersal Rate of FTB", xlab = "Distance Class") 

curve(
  exp_fun(x, 1.5, 0.003),
  from = 0, to = 1500, add = TRUE, lty = "solid", col = "black")

```

Q12 (2 pts): Ricker Model Provide the values of the a and b. Briefly describe how you chose the values.

  a = 0.01 b = 0.005 I don't fully understand these models, so I kept plugging in different values until it looked like it fit my data.

Q13 (2 pts): In your lab report, include a scatterplot of the salamander data with your fitted ricker model.
```{r}
ricker_fun = function(x, a, b) 
{
  return(a * x * exp(-b * x))
}

plot(disp.rate.ftb ~ dist.class, data = sal_dispersal, pch = 20, col = "Blue", 
     main = "Ricker Function of Salamander Dispersal", 
     ylab = "Dispersal Rate of FTB", xlab = "Distance Class")

curve(
  ricker_fun(x, 0.01, 0.005), 
  from = 0, to = 1500, add = TRUE, 
  main = "Ethan's Ricker Function",
  ylab = "frequency of (x)", xlab = "x", lty = "solid", col = "black")

```

Q14. Show the R code you used to create your data frame of model residuals.

Q15. In your lab report, include histograms of the residuals for each of your three models. You may create a single figure with three panels, or include three separate figures.
```{r echo=TRUE, fig.height=9, fig.width=7}
sal_dispersal = data.frame(read.csv(here("data", "dispersal.csv")))

sal_dispersal$y_predicted_linear = line_point_slope(sal_dispersal$dist.class, 800, 0.2, -0.0005)

sal_dispersal$y_predicted_expo = exp_fun(sal_dispersal$dist.class, 1.5, 0.003)

sal_dispersal$y_predicted_rick = ricker_fun(sal_dispersal$dist.class, 0.01, 0.005)


dat_dispersal = data.frame(y_observed = sal_dispersal$disp.rate.ftb)


dat_dispersal$lin_res = dat_dispersal$y_observed - sal_dispersal$y_predicted_linear
dat_dispersal$expo_res = dat_dispersal$y_observed - sal_dispersal$y_predicted_expo
dat_dispersal$rick_res = dat_dispersal$y_observed - sal_dispersal$y_predicted_rick

par(mfrow = c(3, 1))
hist(dat_dispersal$lin_res, main = "Linear Function Residuals", ylab = "Frequency", xlab = "Residuals")
hist(dat_dispersal$expo_res, main = "Exponential Function Residuals", ylab = "Frequency", xlab = "Residuals")
hist(dat_dispersal$rick_res, main = "Ricker Function Residuals", ylab = "Frequency", xlab = "Residuals")


```
