---
title: "Lab 9 Modeling 2"
author: "Ethan Rutledge"
date: "11/21/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Q1 (1 pt.): State the null hypothesis of the Chi-square test.

Brown Creepers will have no difference in presence/absence between edge and interior habitats.

Q2 (2 pts.): Consider the results of your test and explain whether you think that Brown Creepers show a significant habitat preference.

Brown Creepers have a habitat preference; interior habitats, pvalue <.05


```{r}
require(here)
birds   = read.csv(here("data", "bird.sta.csv"))
hab     = read.csv(here("data", "hab.sta.csv"))
birdhab = merge(birds, hab, by=c("basin", "sub", "sta"))

# Create a contingency table for edge/interior and brown creeper presence/absence
table(birdhab$s.edge, birdhab$BRCR > 0)
br_creeper_table = table(birdhab$s.edge, birdhab$BRCR > 0)[, 2:1]
colnames(br_creeper_table) = c("present", "absent")

chisq.test(br_creeper_table)


```

Q3 (1 pt.): Show the R-code you can use to create a model fit (call it fit_species) of penguin body mass as predicted by penguin species.
Q4 (1 pt.): Show the R-code you can use to create a model fit (call it fit_sex) of penguin body mass as predicted by sex.
Q5 (1 pt.): Show the R-code you can use to create a model fit (call it fit_both) of penguin body mass as predicted by species and sex.


```{r}
require(palmerpenguins)

fit_species = 
  lm(
    formula = body_mass_g ~ species,
    data = penguins)

fit_sex = 
  lm(
    formula = body_mass_g ~ sex,
    data = penguins)

fit_both = 
  lm(
    formula = body_mass_g ~ species * sex,
    data = penguins)

```

Q6 (1 pt.): Include a conditional boxplot corresponding to your fit_species model.
Q7 (1 pt.): Include a conditional boxplot corresponding to your fit_sex model.
Q8 (3 pts.): Include a conditional boxplot corresponding to your fit_both model.

```{r}
boxplot(body_mass_g ~ species, data = penguins, 
        ylab = "Body Mass (g)",
        xlab = "Species",
        main = "Penguin Boxplot of \nBody Mass by Species",
        names = c("Adelie", "Chinstrap", "Gentoo"))

boxplot(body_mass_g ~ sex, data = penguins,
        ylab = "Body Mass (g)",
        xlab = "Sex",
        main = "Penguin Boxplot of \nBody Mass by Sex",
        names = c("Female", "Male"))

boxplot(body_mass_g ~ species * sex, data = penguins, las = 2,
        ylab = "Body Mass (g)",
        xlab = "",
        main = "Penguin Boxplot of \nBody Mass by Species/Sex",
        names = c("Adelie \nFemale", "Adelie \nMale", "Chinstrap \nFemale",
"Chinstrap \nMale","Gentoo \nFemale", "Gentoo \nMale"))
```

Q9 (3 pts.): Based on the shapes of the boxes, which of the models (if any) do you think may have problems fulfilling the homogeneity assumption?

They all look like they fulfill the homogeneity assumption because they are all more or less close in size, fit species may be a little off.

Q10 (1 pt.): State the null hypothesis of the Bartlett test.
There will be no differences in variances between the groups.

Q11 (1 pt.): What was the p-value from the Bartlett test of homogeneity for observations grouped by species?

p-value = 0.0501

Q12 (1 pt.): What was the p-value from the Bartlett test of homogeneity for observations grouped by sex?

p-value = 0.03194

```{r}
bartlett.test(body_mass_g ~ species,
    data = penguins)

bartlett.test(body_mass_g ~ sex, 
    data = penguins)
```

Q13 (1 pt.): What was the p-value from the Bartlett test of homogeneity for observations grouped by both factors?

p-value = 0.1741

```{r}
dat_groups = aggregate(
  body_mass_g ~ species * sex,
  data = penguins,
  FUN = c)

dat_groups$body_mass_g

bartlett.test(dat_groups$body_mass_g)

```
Q14 (3 pts.): Based on the results of the Bartlett tests, do you anticipate any issues with heterogeneity in any of the models? Make sure you justify your response with the results of your tests.

For the species model we got a p-value = 0.0501 and for the sex model we got a p-value = 0.03194. Thus we would reject the null hypothesis that is is no difference in variances. So for these two models there are issues with heterogeneity. For the model with both factors the p-value = 0.1741, so we would fail to reject the null hypothesis that variances were different. So there is no issue with heterogeneity for this model. 
