---
title: "Plot and Logistic Curve Code"
author: "Ethan Rutledge"
date: "9/23/2021"
output: pdf_document
---


```{r}
##Install package psych in order to make pair plots
install.packages("psych")
require(psych)

```

```{r}
##bring in data
require(here)
read.csv(here("data", "hab.sta.csv"))

##create data frame
dat_habitat = data.frame(read.csv(here("data", "hab.sta.csv")))

##inspect data with head function
head(dat_habitat)

##bring in data
read.csv(here("data", "bird.sta.csv"))

##create data frame
dat_bird = data.frame(read.csv(here("data", "bird.sta.csv")))

##inspect data with head function
head(dat_bird)

```

```{r, echo=TRUE, width=1600, height=1400}
##change image output^^

##merge 2 data sets with identical variables
dat_all = merge(dat_habitat, dat_bird)

##plot
plot(ba.tot ~ elev, data = dat_all)

##save image to image library to save space
png(filename = here("Images", "Call it whatever.png"))
plot(ba.tot ~ elev, data = dat_all)
dev.off()

```

```{r}
##random sample of 100 variables to check out data
sample(dat_all$CEWA, 50)

```

```{r}
##create boolean vector of  all the times cedar wax wing was observed
cewa_vec = (dat_bird$CEWA > 0)

##change to numeric present absent data
cewa_present_absent = as.numeric(cewa_vec > 0)

##plot it
plot(x = dat_all$elev, y = cewa_present_absent)

```

```{r}
# Function to calculate the logistic parameter a given the slope and midpoint
get_logistic_param_a = function(slope, midpoint)
{b = slope / 4 
  return (-midpoint * (slope / 4))}

# Function to calculate the logistic parameter b given the slope
get_logistic_param_b = function(slope)
  {return (slope / 4)}


# Calculate the value of the logistic function at x, given the parameters a and b.
logistic = function(x, a, b)
{val = exp(a + b * x)
  return(val / (1 + val))}

# Calculate the value of the logistic function at x, given a slope and midpoint.
logistic_midpoint_slope = function(x, midpoint, slope)
{b = get_logistic_param_b(slope)
  a = get_logistic_param_a(slope, midpoint)
  return(logistic(x, a, b))}

```

```{r}
##plot a logistic curve. It looks like a good midpoint value could be 400. We can try a slope of 0.1 to start with
plot(x = dat_all$elev, y = cewa_present_absent)
curve(logistic_midpoint_slope(x, midpoint = 400, slope = 0.1), add = TRUE)

```

```{r}
##try a negative slope
plot(x = dat_all$elev, y = cewa_present_absent)
curve(logistic_midpoint_slope(x, midpoint = 400, slope = -0.1), add = TRUE)

```

```{r}
##try a shallower negative slope
plot(x = dat_all$elev, y = cewa_present_absent)
curve(logistic_midpoint_slope(x, midpoint = 400, slope = -0.05), add = TRUE)

```

```{r}
## make it a pretty plot
plot(x = dat_all$elev, y = cewa_present_absent, col = "blue", cex = 1, pch = 16)
curve(logistic_midpoint_slope(x, midpoint = 400, slope = -0.05), add = TRUE)

```
Assignment: More practice

1. Use the pair plot function from psych to create a pair plot of the three terrain variables and basal area from the lecture questions.



```{r}
## create a pair plot using the 3 terrain variables and basal area, the [,c] refers to using a subset from the dat_all dataset from columns elev,slope,aspect, as opposed to [c,] which refers to all rows.
pairs.panels(dat_all[,c("elev", "slope", "aspect", "ba.tot")])

```

2. Choose two bird species and create plots of presence/absence (on the y-axis) and basal area (on the x axes).

3. Visually inspect the plots and fit logistic curves using the parameterization functions provided above.

```{r}
##create boolean vector of  all the times common yellow throat was observed
coye_vec = (dat_all$COYE > 0)

##change to numeric present absent data
coye_present_absent = as.numeric(coye_vec > 0)

##plot it
plot(x = dat_all$ba.tot, y = coye_present_absent, ylab = "Common Yellow Throat Presence/Absence", xlab = "Total Basal Area")

##fit logistic curve
curve(logistic_midpoint_slope(x, midpoint = 100, slope = -0.5), add = TRUE)

##at very low total basal area we find a few birds, but any thicker and theyre out.

```

```{r}
##create boolean vector of  all the times robin was observed
amro_vec = (dat_all$AMRO > 0)

##change to numeric present absent data
amro_present_absent = as.numeric(amro_vec > 0)

##plot it
plot(x = dat_all$ba.tot, y = amro_present_absent, ylab = "American Robin Presence/Absence", xlab = "Total Basal Area")

##fit logistic curve
curve(logistic_midpoint_slope(x, midpoint = 100, slope = -0.5), add = TRUE)

##we see the presence of robins matches the absence of them up to 100m per ha, looks like a 50% chance of seeing them up to that point.
```

How many total number of Gray Jays were observed in all of the sampling sites.
Try using the dat_all data frame to calculate the answer

```{r}
sum(dat_all$GRJA)

```

Calculate the total number of sampling sites in which Gray Jays were observed.
Hint: What happens when you use the sum() function on a vector of Boolean values?

```{r}
grja_vec = (dat_all$GRJA > 0)
grja_present_absent = as.numeric(grja_vec > 0)
sum(grja_present_absent)

```


.