---
title: "Final Project R Reference"
author: "Ethan Rutledge"
date: "12/2/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# R Reference Guide {.tabset .tabset-pills}


## Loading Data and Packages


### Load the (here) and (palmerpenguins) packages using library() and require() and use the 2021 ginkgo data to create a data.frame called ginkgo.


```{r}

##bring in data
require(here)

##library does the same thing but takes longer
library(palmerpenguins)

## create data frame
ginkgo = data.frame(read.csv(here("data", "ginkgo_data_2021.csv")))

## preview data
head(ginkgo)


```



## Data Structures


### c()


The function c() combines or concatenates its arguments into a vector (a 1-dimensional data structure consisting of 1 or more elements).

* All of the elements must be of the same type.
    * I can’t combine character and numeric types in the same call to c()

Here’s two examples using numeric and character data types:


```{r}

## Create a vector of numbers:
num_vec  = c(1, 4, 8, 9, 13)

## Create a vector of characters:
char_vec = c("a", "fish", "data is cool")

## I can show the contents of a vector by typing the name of the vector, or using the print() function.

## Typing the name of the vector into the console prints the contents
num_vec

## The print() function accomplishes the same task:
print(char_vec)

```


### length


The function length() function gets or sets the length of a vector (list) or other objects. ... length() function can be used for all R objects. For an environment it returns the object number in it.

Returns the number of characters in a string.


```{r}
length(num_vec)

sum(num_vec)
```
### matrix

matrix creates a matrix from the given set of values.

```{r}
matrix(num_vec)
```

### nrow(), ncol() and dim()


```{r}

ginkgo_1 = data.frame(read.csv(here("data", "ginkgo_data_2021.csv")))

##the Number of Rows/Columns of an Array
nrow(ginkgo_1)
ncol(ginkgo_1)

##dimensions of an object
dim(ginkgo_1)

```


## Subsetting

### $ 

### [] 

### subset()


```{r}

##One column
ginkgo_seeds = (ginkgo_1$seeds_present)

ginkgo_seeds

##select first row of the ginkgo data
ginkgo_row_1 = ginkgo_1[1,]

ginkgo_row_1

##select the element in row 2, column 3
ginkgo_row2_col3 = ginkgo_1[2,3]

ginkgo_row2_col3

##select the 3rd column of the ginkgo data
ginkgo_col3 = ginkgo_1[,3]

ginkgo_col3

##retrieve all the data for Adelie penguins.
dat_pen = (subset(penguins, species = "Adelie"))

dat_pen

##leave out data for Adelie penguins
dat_pen_1 = (subset(penguins, species != "Adelie"))

dat_pen_1

```


## Numerical Data Exploration

### Summary, Mean, SD


```{r}

summary(ginkgo_1)
mean(ginkgo_1$max_depth)
sd(ginkgo_1$max_depth)

```



## Graphical Data Exploration


### Scatterplot, Histogram, Boxplot 


```{r}

plot(max_depth ~ max_width, data = ginkgo_1,
  col = 2,
  pch = 20,
  cex = 1,
  main = "Gingko Depth by Width",
  xlab = "Depth",
  ylab = "Width",
  xlim = c(0,150),
  ylim = c(0,150)
    )


hist(penguins$flipper_length_mm, breaks = 12)


boxplot(ginkgo_1$petiole_length)

boxplot(max_depth ~ seeds_present, data = ginkgo_1)

```
### 4-panel figure of histograms, arranged in a 2 by 2 grid. You may use any data you like, but each histogram must be different and have appropriate titles and axes. - par() required arguments: - mrow =

```{r}
par(mfrow = c(2, 2))
hist(penguins$bill_depth_mm, main = "Penguin Bills", ylab = "Frequency", xlab = "Bill Depth")
hist(ginkgo_1$max_depth, main = "Ginkgo Leaf", ylab = "Frequency", xlab = "Leaf Depth")
hist(penguins$flipper_length_mm, main = "Penguin Flippers", ylab = "Frequency", xlab = "Flipper Length")
hist(penguins$body_mass_g, main = "Penguin Weight", ylab = "Frequency", xlab = "Body Mass")

```


## Distribution Functions

### dnorm()

### pnorm()

### qnorm()

### dbinom()

### pbinom()

### qbinom()

What is the probability of observing a count of exactly 3 successes in a binomial distribution with parameters n = 4 and p = 0.75?

```{r}

dbinom(3, 4, 0.75)

```


What is the probability of observing a count of 3 successes or fewer in a binomial distribution with parameters n = 4 and p = 0.75?

```{r}
pbinom(3, 4, 0.75)

```

What is the 90th percentile?: The quantile function, e.g. qbinom()

```{r}
qbinom(3, 4, 0.75)

```


What is the probability of observing more than 3 successes in a binomial distribution with parameters n = 5 and p = 0.75?

```{r}
1 - pbinom(3, 5, 0.75)

```

The probability density function, e.g. dnorm()

```{r}
dnorm(2.4, 2, 2)

```

What is the probability of observing a value of less than 1.2 from a normally-distributed population with mean = 2 and standard deviation = 2?

```{r}
pnorm(1.2, mean = 2, sd = 2)

```


What is the probability of observing a value of greater than 1.2 from a normally-distributed population with mean = 2 and standard deviation = 2?

```{r}
1 - pnorm(1.2, mean = 2, sd = 2)

```


 What is the probability of observing a value between 1.2 and 3.2 from a normally-distributed population with mean = 2 and standard deviation = 2?

```{r}
pnorm(3.2, mean = 2, sd = 2) - pnorm(1.2, mean = 2, sd = 2)

```


What is the 20th percentile of flipper lengths?: The quantile function, e.g. qnorm()

```{r}
qnorm(3.2, 2, 2)

```

























































