---
title: "InClass"
author: "Ethan Rutledge"
date: "11/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
require(palmerpenguins)

```

```{r}

dat_ade = droplevels(subset(penguins, species == "Adelie"))

hist(dat_ade$body_mass_g, main = "Adelie Penguins: Body Mass", xlab = "body mass (g)")
```

Create separate boxplots of body mass for male and female Adelie penguins. Your boxplots do not have to be in the same panel like mine. Show the R-code you used to make the plots.

```{r}
dat_mass = droplevels(subset(penguins, species == "Adelie"))

boxplot(body_mass_g ~ sex, data = dat_mass, ylab = "Body Mass g", xlab = "Sex", main = "Body Mass by Sex of Adelie Penguins")

```

Perform a one-sample t-test of the alternative hypothesis that female Adelie penguins have a body mass different from zero grams. Note that this is a very silly alternative hypothesis. Is this a one- or two-tailed test? Show your R-code.

Yes, so we reject the null hypothesis that there is no difference between females and 0g
```{r}
dat_gurl = (subset(dat_mass, sex == "female"))


t.test(dat_gurl$body_mass_g,
       alternative = c("two.sided"),
       mu = 0)



```

Now, conduct a slightly less silly test: perform a one-sample t-test of the null hypothesis that male Adelie penguins have a mean body mass greater than 4000 grams. Is this a one- or two-tailed test?

No, its less so we fail to reject the null
```{r}

dat_guy = (subset(dat_mass, sex == "male"))

t.test(dat_gurl$body_mass_g,
       alternative = c("greater"),
       mu = 4000)
```
Conduct a two-sample t-test of the alternative hypothesis that male and female Adelie penguins have different mean body masses. Show your r-code.

```{r}
t.test(dat_gurl$body_mass_g, dat_guy$body_mass_g)


```

Conduct a two-sample t-test of the alternative hypothesis that male and female Adelie penguins have different mean body masses. Show your r-code. (Same thing but easier)
```{r}

t.test(body_mass_g ~ sex, data = dat_ade)

```

Conduct a two-sample (one-tailed) t-test of the directional alternative hypothesis that male Adelie penguins are heavier than females. They Are

```{r}
t.test(dat_guy$body_mass_g, dat_gurl$body_mass_g, alternative = "greater")

```


Conduct a two-sample (one-tailed) t-test of the directional alternative hypothesis that female Adelie penguins are heavier than males. They are not.

```{r}
t.test(dat_gurl$body_mass_g, dat_guy$body_mass_g, alternative = "greater")

```






