---
title: "Using_Models_2"
author: "Ethan Rutledge"
date: "11/27/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
require(palmerpenguins)

```

Q1 (4 pts.): Re-create the conditional boxplot of penguin body mass conditioned on sex and species.

Q2 (2 pts.): Based on the boxplots, do you think male penguins (of any species) are significantly heavier than female penguins? Explain your reasoning.
Yes, Gentoo males appear to be significantly heaver than Gentoo females. The error bars/confidence intervals do not overlap.

Q3 (2 pts.): Do you think adding sex to a model that already includes species will improve the model fit?

Make sure you justify your answer based on the boxplots and not results of a statistical test.

I think it adds more evidence that there are differences in body mass between species, but we may arrive at that conclusion without sex.


```{r}

boxplot(body_mass_g ~ species * sex, data = penguins, las = 2,
        ylab = "Body Mass (g)",
        xlab = "",
        main = "Penguin Boxplot of \nBody Mass by Species/Sex",
        names = c("Adelie \nFemale", "Adelie \nMale", "Chinstrap \nFemale",
"Chinstrap \nMale","Gentoo \nFemale", "Gentoo \nMale"))

```

Q4 (2 pts.): Show the R-code you used to build fit_both.

Q5 (2 pts.): What is the base case for the two-way model that includes sex and species?
sexfemale

```{r}
fit_both = lm(body_mass_g ~ sex * species, data = penguins)

summary(fit_both)

anova(fit_both)

```


Q6 (2 pts.): What are the names of the two coefficients (from the first column of the coefficient table) you need to calculate the average mass of female Chinstrap penguins?
sexfemale and specieschinstrap

Q7 (2 pts.): What is the predicted average mass of female Chinstrap penguins in the interactive model?
3527.21

Q8 (2 pts.): What is the observed average mass of female Chinstrap penguins, calcluated from the penguins data?
3527.206

```{r}
aggregate(body_mass_g ~ species * sex, data = penguins, FUN = mean)

```






























