---
title: "Using_Models_2"
author: "Ethan Rutledge"
date: "11/27/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


1-sample t-test on the Gentoo penguin flipper lengths = 0
```{r}
require(palmerpenguins)

t.test(subset(penguins, species == "Gentoo")$flipper_length_mm)

```
1-sample t-test on the Gentoo penguin flipper lengths = 218
```{r}
t.test(
  x = subset(penguins, species == "Gentoo")$flipper_length_mm,
  mu = 218
)

```
one-tailed alternative hypothesis: Gentoo penguin flippers are smaller than 218 mm
```{r}
t.test(
  x = subset(penguins, species == "Gentoo")$flipper_length_mm,
  mu = 218,
  alternative = "less"
)

```
2 sample Ttest compare the flipper lengths of two species
```{r}
t.test(flipper_length_mm ~ species, data = subset(penguins, species != "Chinstrap"))

```

Look at the data

```{r}
par(mfrow = c(1, 2))
hist(penguins$body_mass_g, breaks = 80, main = "histogram of body mass", xlab = "body mass (g)")
plot(density(penguins$body_mass_g, na.rm = TRUE), main = "density plot of body mass")

```
Look at the data
```{r}

boxplot(body_mass_g ~ species, data = penguins)

```
look at the means
```{r}

dat_chinstrap = subset(penguins, species == "Chinstrap")
mean(dat_chinstrap$body_mass_g, na.rm = TRUE)

```
test whether within-group data are normally-distributed:
```{r}
shapiro.test(dat_chinstrap$body_mass_g)

```

Here’s a cool shortcut for calculating the species mean body masses using aggregate() and the formula notation
```{r}
aggregate(body_mass_g ~ species, data = penguins, FUN = mean)

```

Same thing with the shapiro test

```{r}
aggregate(
  body_mass_g ~ species,
  data = penguins,
  FUN = function(x) shapiro.test(x)$p.value)

```

Fit a linear model,Then we can look at the model coefficients:

```{r}
fit_species = lm(body_mass_g ~ species, data = penguins)
summary(fit_species)

```
ANOVA
```{r}
anova(fit_species)

```

The syntax to build the model in R is easy. Use the lm() function to store the model in a variable: ANOVA complete walkthrough

```{r}
fit_species = lm(body_mass_g ~ species, data = penguins)

summary(fit_species)

anova(fit_species)

```

2 Way ANOVA, without interaction terms

```{r}
fit_additive = lm(body_mass_g ~ sex + species, data = penguins)

```

2 Way Anova, with interaction terms
```{r}
fit_interactive = lm(body_mass_g ~ sex * species, data = penguins)

summary(fit_interactive)

anova(fit_interactive)

```

Simple Linear Regression

```{r}
lm(bill_length_mm ~ body_mass_g, data = penguins)


```





























