---
title: "Final Project R Reference"
author: "Ethan Rutledge"
date: "12/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# R Reference Guide {.tabset .tabset-pills}


## Loading Data and Packages


Show how to load the (here) and (palmerpenguins) packages - library() and require()

Use the 2021 ginkgo data to create a data.frame called ginkgo.


```{r}

##bring in data
require(here)

##library does the same thing but takes longer
library(palmerpenguins)

## create data frame
gingko = data.frame(read.csv(here("data", "ginkgo_data_2021.csv")))

## preview data
head(gingko)


```



## Data Structures


### c()


The function c() combines or concatenates its arguments into a vector (a 1-dimensional data structure consisting of 1 or more elements).

* All of the elements must be of the same type.
    * I can’t combine character and numeric types in the same call to c()

Here’s two examples using numeric and character data types:


```{r}

## Create a vector of numbers:
num_vec  = c(1, 4, 8, 9, 13)

## Create a vector of characters:
char_vec = c("a", "fish", "data is cool")

## I can show the contents of a vector by typing the name of the vector, or using the print() function.

## Typing the name of the vector into the console prints the contents
num_vec

## The print() function accomplishes the same task:
print(char_vec)

```


### length


The function length() function gets or sets the length of a vector (list) or other objects. ... length() function can be used for all R objects. For an environment it returns the object number in it.

Returns the number of characters in a string.


```{r}


```



## Subsetting


## Numerical Data Exploration


## Graphical Data Exploration


## Distribution Functions





























































